---
title: "Black Friday Sales"
author: "mihir"
date: "7 October 2017"
output: html_document
---
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(dummies)
```


#Let us load the dataset 


```{r}
training<-read.csv('C://Users//MD//Desktop//PROJECTS//blackfriday//train.csv')
testing<-read.csv('C://Users//MD//Desktop//PROJECTS//blackfriday//test.csv')
```

#Now let us look at some summary statistics:-

```{r}
summary(training)
str(training)
View(training)
```

For Further data processing and EDA,we will combine training and testing datasets 

```{r}
testing$Purchase<-0
combinebf<-rbind(training,testing)
```

We will convert categorical data into factors

```{r}
combinebf$Occupation<-as.factor(combinebf$Occupation)
combinebf$Product_Category_1<-as.factor(combinebf$Product_Category_1)
combinebf$Product_Category_2<-as.factor(combinebf$Product_Category_2)
combinebf$Product_Category_3<-as.factor(combinebf$Product_Category_3)
```

#Let's do some plots

Plot of Purchase against city type wrt gender:-

```{r}
ggplot(training, aes(x = City_Category, y = Purchase, fill = Gender)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("City") + 
  ylab("Purchase") + 
  ggtitle("Purcahse vs City")
```

Plot of purchase against Age group wrt Gender:-

```{r}
ggplot(training, aes(x = Age, y = Purchase, fill = Gender)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Age") + 
  ylab("Purchase") + 
  ggtitle("Purchase vs Age")
```

Plot Occupation Vs Product Category:-

```{r}
ggplot(combinebf[1:nrow(training),], aes(x = Occupation, y =Purchase,fill=Gender))+
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Occupation") + 
  ylab("Purchase") + 
  ggtitle("Occupation vs Purchase")
```

Lets plot the histogram for Purchase:-

```{r}
ggplot(data = training,aes(x=Purchase))+geom_histogram(binwidth =500)+ggtitle("Histogram of Purchase")
plot(density(training$Purchase))
```


It's nearly normally distributed with a little skew towards the right.

Let's check which variable has NA's-

```{r}
sapply(combinebf,function(x) sum(is.na(x)))
```

Product Category 2 and 3 have high number of NA's.We will have to treat it.

We will substitute the NA with some random large number:-
```{r}
combinebf$Product_Category_2<-as.integer(combinebf$Product_Category_2)
combinebf$Product_Category_2[is.na(combinebf$Product_Category_2)]<-9999
combinebf$Product_Category_3<-as.integer(combinebf$Product_Category_3)
combinebf$Product_Category_3[is.na(combinebf$Product_Category_3)]<-999
```


#Let's do some data cleaning:-


We will encode city variable:-

```{r}
combinebf <- dummy.data.frame(combinebf, names=c("City_Category"), sep="_")
```

We will convert age variable to numeric  from factor-

```{r}
combinebf$Age<-as.character(combinebf$Age)
combinebf$Age[combinebf$Age == "0-17"] <- "15"
combinebf$Age[combinebf$Age == "18-25"] <- "21"
combinebf$Age[combinebf$Age == "26-35"] <- "30"
combinebf$Age[combinebf$Age == "36-45"] <- "40"
combinebf$Age[combinebf$Age == "46-50"] <- "48"
combinebf$Age[combinebf$Age == "51-55"] <- "53"
combinebf$Age[combinebf$Age == "55+"] <- "60"

```

```{r}
combinebf$Age<-as.integer(combinebf$Age)
```

Converting stay in current city to numeric:-
```{r}
combinebf$Stay_In_Current_City_Years<-as.character(combinebf$Stay_In_Current_City_Years)
combinebf$Stay_In_Current_City_Years[combinebf$Stay_In_Current_City_Years=="4+"]<-"4"
combinebf$Stay_In_Current_City_Years<-as.integer(combinebf$Stay_In_Current_City_Years)
```

We will convert the Gender Variable to binary:-


```{r}
combinebf$Gender<-ifelse(combinebf$Gender=="M",0,1)
```

We will create new features:-

Count of each product:-
```{r}
product_count <- ddply(combinebf, .(Product_ID), nrow)
names(product_count)[2] <- "Product_Count"
combinebf<-merge(combinebf,product_count,by="Product_ID")
combinebf$Product_Count[is.na(combinebf$Product_Count)]<-0
```

Average purchase amount for each feature:-

```{r}
product_mean <- ddply(combinebf[1:nrow(training),], .(Product_ID), summarize, Product_Mean=mean(Purchase))
combinebf<-merge(combinebf,product_mean,by="Product_ID",all = TRUE)
combinebf$Product_Mean[is.na(combinebf$Product_Mean)]<-0
```

User Count:-
```{r}
usercount <- ddply(combinebf, .(User_ID), nrow)
names(usercount)[2] <- "User_Count"
combinebf<-merge(combinebf,usercount,by="User_ID",all = TRUE)
```


Lets remove some variables from our dataframe

```{r}
combinebf1<-subset(combinebf,select = -c(Product_ID,User_ID))
```

Now lets split the dataset again into test and train for our modelling:-

```{r}
train<-combinebf1[1:nrow(training),]
test<-combinebf1[(nrow(training)+1):nrow(combinebf1),]

```

Let's start with our modelling:-




```{r}
library(caret)
```


Using gradient boosting for Predictions:-

```{r}
TrainControl <- trainControl( method = "repeatedcv", number = 4, repeats = 3)
model<- train(Purchase~ ., data = train, method = "gbm", trControl = TrainControl,verbose = FALSE)

```

```{r}
print(model)
```




```{r}
test$predicted<-predict(model,test)
```

```{r}
print(head(test$predicted,10))
```


```{r}
bfdf<-testing%>%select(User_ID,Product_ID)
bfdf1<-cbind(bfdf,test$predicted)
```

We will now write the predictions on disc:-

```{r}
write.csv(bfdf1,file = "C://Users//MD//Desktop//PROJECTS//blackfriday//submission.csv")
```

After Submitting the csv on AV portal the RMSE for the solution was a  2936.74 which has not far from the best solution of around 2450.


